Cost- and Deadline-Constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-04-23 paper/paper.tex
Remove paragraphs that were eliminated

\begin{figure}[tb] 
  \centering
  \includegraphics[width=\textwidth]{figures/spss-gantt}
  \caption[Example schedule generated by the SPSS algorithm]{Example schedule generated by the SPSS algorithm. Each row is a different VM. Tasks are represented as boxes and are colored by workflow.}
  \label{fig:spss-example}
\end{figure}

@@

\subsection{Scenarios}
\label{sec:scenarios}

\TODO{Edit this to remove duplication with section \ref{sec:ensembles}}

\begin{figure*}[t] 
\centering
\includegraphics[width=0.19\textwidth]{pareto-MONTAGE-n-1000-8-dagh1-20m0}
\includegraphics[width=0.19\textwidth]{pareto-CYBERSHAKE-n-1000-8-dagh1-20m0}
\includegraphics[width=0.19\textwidth]{pareto-LIGO-n-1000-8-dagh1-40m0}
\includegraphics[width=0.19\textwidth]{pareto-GENOME-n-1000-8-dagh100-1500m0}
\includegraphics[width=0.19\textwidth]{pareto-SIPHT-n-1000-8-dagh5-50m0}
\caption{Number of workflows completed for DPDS, AW-DPDS and SPSS
algorithms on ensembles of 100 Pareto-distributed workflows, {\em small budget}
(top) and {\em large budget} (bottom). The ranges on $y$ axis are chosen to 
emphasize the differences between the performance of the algorithms.}
\label{fig:number-complete-pareto}
\end{figure*}

The number of workflows in an ensemble depends on the particular application, but
we assume that ensemble sizes are on the order of between 10 and 100 workflows.
This range is motivated by several reasons: First, such sizes are typical of the
real applications we have examined. For example, the number of geographical 
sites of interest to the users of the CyberShake application is on the order 
of 100. Second, smaller ensembles consisting of just a few workflows can be 
aggregated into a single workflow, so there is no need to treat them as an
ensemble. Similarly, when the number of workflows grows, and each 
workflow has a large number of tasks, either the constraints are low enough to
prevent many of the workflows from running, or the problem of efficiently 
allocating them to the resources becomes similar to the bag-of-tasks problem, 
which is easier to solve efficiently.

In order to evaluate the algorithms on a standard set of workflows, we used
ensembles constructed from the workflows available\footnote{\url{https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator}}
in the workflow generator gallery~\cite{Bharathi2008}. The gallery contains synthetic
workflows modeled on real applications. We used workflows from five applications 
that were managed by the Pegasus Workflow Management System. For each
application we constructed ensembles of 100 workflows by sampling collections of
workflows of various sizes (number of tasks) based on a given distribution.
Two distributions are considered:

\begin{itemize}
  \item constant distribution
  \item Pareto-like distribution.
\end{itemize}

The constant distribution produces ensembles containing workflows with the same number of tasks per workflow, 
while the Pareto-like distribution produces ensembles with a small number of larger 
workflows and a large number of smaller workflows, as seen in 
Figure~\ref{fig:ensemble-pareto}. The number of large workflows (of size $\geq$ 900) 
is slightly increased to represent the ``heavy-tail'', i.e. that there is a 
significant number of large workflows in the ensemble. We assume that large
workflows are more important to users, so in the Pareto ensembles the workflows are 
assigned priorities according to their size so that the largest workflows are given 
the highest priority. In
future work we plan to explore other ensemble types such as uniform distributions.

\begin{figure*}[t]  
\centering
\includegraphics[width=0.19\textwidth]{pareto-size-MONTAGE-n-1000-8-dagh1-20m0}
\includegraphics[width=0.19\textwidth]{pareto-size-CYBERSHAKE-n-1000-8-dagh1-20m0}
\includegraphics[width=0.19\textwidth]{pareto-size-LIGO-n-1000-8-dagh1-40m0}
\includegraphics[width=0.19\textwidth]{pareto-size-GENOME-n-1000-8-dagh100-1500m0}
\includegraphics[width=0.19\textwidth]{pareto-size-SIPHT-n-1000-8-dagh5-50m0}
\caption{ Total work completed expressed as the sum of runtimes (in hours) of
tasks of completed workflows. Data computed for DPDS, AW-DPDS and SPSS
algorithms on ensembles of 100 Pareto-distributed workflows. The ranges on $y$ axis are chosen to 
emphasize the differences between the performance of the algorithms.}
\label{fig:total-time}
\end{figure*}

In the experiments we assumed that all the VMs are homogeneous, have a
processing speed of 1000 MIPS (million instructions per second), and a price of
\$1 per VM-hour. We also assume that the sizes of tasks in the workflow gallery
are given in seconds measured on a 1000 MIPS machine. These values were chosen
to approximate realistic numbers, and also to simplify interpretation of
results. Moreover, the absolute values of these parameters do not affect the
relative comparison between the algorithms. In this study we do not take into
account the heterogeneity of the infrastructure since we assume that it is
always possible to select a VM type that has the best price to performance ratio
for a given application.

For each application, we selected ranges of constraints (deadline and budget) so
that it is possible to observe the possibly broad result space: from tight 
constraints, where only a small number of workflows can be completed, to 
more liberal constraints where all, or almost all, of the workflows can be 
completed. These parameters vary among the applications. Montage and CyberShake 
ensembles required budgets in the \$20 -- \$150 range and deadlines from 1 to
20 hours to be completed. The LIGO and SIPHT workflows have more long-running
tasks, so their ensembles required budgets up to \$1,500 and deadlines up to 
40 and 60 hours. The Epigenomics workflow is even more costly, requiring 
approximately \$12,000 and 1,500 hours to complete a full ensemble. These
differences in total runtime and cost come from the fact that we used the
same workflow size distribution for all five applications, but
the sizes of tasks vary between applications.

We show the results for three algorithms: DPDS, WA-DPDS and SPSS. These 
experiments were run with maximum autoscaling factor set to 1.0 for
DPDS and WA-DPDS. After experimenting with DPDS and WA-DPDS we found that, 
due to the high parallelism of workflows, the resource utilization remains 
high most of the time, which leads to the rapid provisioning of new resources 
at the beginning of ensemble execution. This, in turn, results in less 
efficient total utilization due to high levels of parallelism. Setting the 
scaling limit to 1.0, ensures that the shape of the resource provisioning 
plan  is close to a rectangle (see Figure~\ref{fig:spds-example}) and the VM 
termination procedure (Algorithm~\ref{alg:prov}) allows maximum utilization 
of VMs before shutdown. Based on experiments with different applications, 
we set the $\alpha$ parameter for deadline distribution in SPSS to be 0.7.

\subsection{Performance Metric}
\label{sec:perf_metric}

@@

There are many different ways to evaluate and compare the performance of 
the algorithms. The set of ensemble performance metrics we have used to
compare our algorithms includes:

@@

The number of workflows completed is a simple metric that can be used for
comparison. However, because our ensembles contain workflows with different 
priorities and sizes, a less efficient algorithm may schedule smaller
workflows first, completing a large number of low-priority workflows. It may 
also happen that a conservative admission algorithm may overestimate the cost 
of a large workflow, thus rejecting high-priority workflows which actually 
could fit into the schedule. Therefore, we defined the user preference score as:

\begin{equation}
\label{eq:score}
s = \sum_{workflows\ completed}{2^{-p}}
\end{equation}

where $p$ is the priority of the workflow such that the highest-priority workflow 
has $p=0$, the next highest workflow has $p=1$, and so on. This exponential scoring 
function gives the highest priority workflow a score that is higher than all the 
lower-priority workflows combined:

\begin{equation}
\label{eq:score-property}
2^{-p} > \sum_{i=p+1,\ldots}2^{-i}
\end{equation}

This scoring is consistent with our definition of the problem, which is to 
complete as many workflows as possible, according to their priorities, given 
a budget and deadline.

@@

@@

We ran simulations for constraint values ranging between the estimated 
limits indicated in Section~\ref{sec:scenarios}, resulting in approximately 
50-150 (budget, deadline) data points for each application. The runs were
performed for Pareto-distributed and constant workflow ensembles. From this result
space we selected representative plots.

\begin{figure*}[htb] 
\centering
\includegraphics[width=0.19\textwidth]{pareto-cost-MONTAGE-n-1000-8-dagh1-20m0}
\includegraphics[width=0.19\textwidth]{pareto-cost-CYBERSHAKE-n-1000-8-dagh1-20m0}
\includegraphics[width=0.19\textwidth]{pareto-cost-LIGO-n-1000-8-dagh1-40m0}
\includegraphics[width=0.19\textwidth]{pareto-cost-GENOME-n-1000-8-dagh100-1500m0}
\includegraphics[width=0.19\textwidth]{pareto-cost-SIPHT-n-1000-8-dagh5-50m0}
\caption{Effective computing cost in \$ per hour calculated by dividing the sum of
task runtimes by the total budget. Higher costs result from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows. The ranges on $y$ axis are chosen to 
emphasize the differences  between  the performance of the algorithms.}
\label{fig:cost}
\end{figure*}

Figure~\ref{fig:number-complete-pareto} shows the number of the workflows
completed for two budget values for each application ensemble. The budget values
selected correspond to two extreme cases: {\em small budget} (top) and {\em
large budget} (bottom). The small budgets permit the algorithms to complete 
only a small number of workflows from the ensemble, while the larger budgets 
enable the algorithms to finish almost all the workflows. In three of
five cases the SPSS algorithm performs better for tighter constraints. When the 
budget is low, or the deadline is short, the static procedure is able to more 
efficiently pack workflow tasks onto limited resources. This is the case for 
the LIGO, Epigenomics and SIPHT ensembles for small budgets, and for large
budgets with short deadlines. The exceptions are the Montage and CyberShake 
ensembles,  which consist of many fine-grained tasks. In these applications, 
as well for the other ensembles with large budgets and deadlines, the simple 
dynamic scheduling techniques perform equally well or better. The reason for 
this behavior is that the large number of tasks guarantees that resources 
are almost never idle regardless of how the tasks are scheduled. We also 
observe that WA-DPDS performs better than the workflow-unaware version, DPDS. 
This means that for the workflows we tested the simple admission procedure based 
on estimation of workflow cost does not degrade the solution, but rather in 
many cases it enables larger workflows that would lead to budget overrun to
be rejected. Thus it can save the space for smaller workflows to complete.

%Figure~\ref{fig:score} shows the score $s$ as defined in Eq.~\ref{eq:score}
%computed for two sample applications. Comparing
%Figure~\ref{fig:number-complete-pareto} to Figure~\ref{fig:score} we can see that
%the latter one gives more flattened shape: this means that e.g. the improvement
%of the solution by one workflow may mean adding a one low-priority workflow.
%Nevertheless we can see that the curves on both figures have a very similar
%shape, which mens that our algorithms obey the priority rules and do not try to
%admit lower-priority workflows at the cost of the more important ones.

\begin{table}[tb]
\centering

(a) Number of workflows completed $N_c$:
\medskip
\begin{tabular}{r|cccccc}
 & M & C & L & E & S & All\tabularnewline
\hline
DPDS      &    55 &  47  & 23  & 14  &  0 & 139\tabularnewline 
WA-DPDS   &    97 &  89  & 50  & 34  &  1 & 271\tabularnewline 
SPSS      &    38 &  37  & 190 & 74  & 50 & 389\tabularnewline
\end{tabular}
\medskip

(b) Total amount of computing time completed $T_C$:
\medskip
\begin{tabular}{r|cccccc}
 & M & C & L & E & S & All\tabularnewline
\hline
DPDS      &   55  & 47  &  26  & 14  &  0 & 142\tabularnewline
WA-DPDS   &   96  & 98  &  59  & 26  & 10 & 289\tabularnewline
SPSS      &   32  & 12  & 165  & 75  & 41 & 325\tabularnewline
\end{tabular}
\medskip

(c) Exponential score $s$:
\medskip
\begin{tabular}{r|cccccc}
 & M & C & L & E & S & All\tabularnewline
\hline
DPDS      &   55  & 47  &  26  & 14  &  0 & 142\tabularnewline
WA-DPDS   &   96  & 98  &  59  & 26  & 10 & 289\tabularnewline
SPSS      &   32  & 12  & 165  & 75  & 41 & 325\tabularnewline
\end{tabular}
\medskip

\caption{Summary table for Pareto-distributed workflows showing number of times
each algorithm achieved the highest score over all budgets and deadlines. Data from a total of 525 runs on 
Pareto-distributed ensembles for Montage (M), CyberShake (C), LIGO (L), 
Epigenomics (E), SIPHT (S).
\label{tab:num-dags-pareto}}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}[tb]
\centering

(a) Number of workflows completed $N_c$:
\medskip
\begin{tabular}{r|cccccc}
 & M & C & L & E & S & All\tabularnewline
\hline
DPDS      &   79  & 92  & 170 &  24 &  29 & 394\tabularnewline
WA-DPDS   &   82  & 100 & 180 &  58 &  27 & 447\tabularnewline
SPSS      &   72  & 77  & 195 &  75 &  50 & 469 \tabularnewline
\end{tabular}
\medskip

(b) Total amount of computing time completed $T_C$:
\medskip
\begin{tabular}{r|cccccc}
 & M & C & L & E & S & All\tabularnewline
\hline
DPDS      &   79  & 92  & 170 &  24 &  29 & 394\tabularnewline
WA-DPDS   &   82  & 100 & 180 &  46 &  27 & 435\tabularnewline
SPSS      &   72  & 75  & 195 &  75 &  50 & 467 \tabularnewline
\end{tabular}
\medskip

(c) Exponential score $s$:
\medskip
\begin{tabular}{r|cccccc}
 & M & C & L & E & S & All\tabularnewline
\hline
DPDS      &   79  & 92  & 170 &  24 &  29 & 394\tabularnewline
WA-DPDS   &   82  & 100 & 180 &  46 &  27 & 435\tabularnewline
SPSS      &   72  & 75  & 195 &  75 &  50 & 467 \tabularnewline
\end{tabular}
\medskip

\caption{Summary table for equal size workflows showing number of times each
algorithm achieved the highest score over all budgets and deadlines. Data from a total of 525 runs on 
ensembles of equal size workflows for Montage (M), CyberShake (C), LIGO (L), 
Epigenomics (E), SIPHT (S).
\label{tab:num-dags-equal}}
\end{table}

Figure~\ref{fig:total-time} presents the same results using the total computing
time metric, $T_C$. These plots confirm our previous observations about the
algorithms' performance, but they also allow us observe a more general property:
for a given budget, when the deadline is tightened the amount of work done
steeply decreases. This is partially due to the fact that fewer workflows can be
completed given a short deadline because their critical paths are simply too long
and they are squeezed out of early start times by high priority workflows.
In addition, assigning a shorter deadline means that more VMs need to be
provisioned to complete the work. However, because of the dependencies in the 
workflows, many of these VMs will either be idle, or will be backfilled with 
tasks from low-priority workflows that never finish, resulting, effectively, 
in reduced resource utilization and fewer completed workflows.

@@

% \begin{figure*}[htb] \centering
% \includegraphics[width=0.19\textwidth]{score2-MONTAGE-n-1000-8-dagh1-20m0}
% \includegraphics[width=0.19\textwidth]{score2-CYBERSHAKE-n-1000-8-dagh1-20m0}
% \includegraphics[width=0.19\textwidth]{score2-LIGO-n-1000-8-dagh1-40m0}
% \includegraphics[width=0.19\textwidth]{score2-GENOME-n-1000-8-dagh100-1500m0}
% \includegraphics[width=0.19\textwidth]{score2-SIPHT-n-1000-8-dagh5-50m0}
% \caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
% Pareto-distributed workflows. Score is computed as $\sum(2^(-p))$ where
% priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
% priority).} \label{fig:} \end{figure*}

\begin{figure}[htb] \centering
\includegraphics[width=0.7\columnwidth]{pareto-budget-SIPHT-n-1000-8-dagh5-50h4m0}\\
(a) ensemble of Pareto-distributed workflows
\includegraphics[width=0.7\columnwidth]{constant-budget-SIPHT-n-1000-0-dagh5-50h4m0}\\
(b) ensemble of workflows of equal size \caption{Comparison of number of
workflows completed for a given deadline, plotted versus different budget
amounts.}
\label{fig:budgets}
\end{figure}

In order to compare the efficiency of our algorithms, we counted the number of times
each algorithm achieved the best result for all simulated applications, deadlines
and budgets. Table~\ref{tab:num-dags-pareto} shows these summary values for 
Pareto-distri\-buted workflows. Note that all algorithms that achieve the best 
result are credited, so the sum for a given application can be greater than the
total number of scenarios. These results depend on the metric: number of
workflows completed does not count the priorities or sizes of workflows, which
means that it happens more often that all the algorithms achieve the same best
result. However, it does not mean that all algorithms complete the same
workflows: that is why the values in (a), (b) and (c) are different. The 
exponential score and total amount of computing time completed give the same
rankings, since in our test ensembles larger workflows were given higher
priorities. It can be observed based on the sampled problem space that the
workflow-aware algorithms achieve the highest score in $\sim$2--3 times more 
cases than the simple DPDS algorithm, which does not take into account workflow 
structure or size. SPSS performs the best in $\sim$20\% more cases than WA-DPDS,
including the Montage and CyberShake ensembles, for which SPSS does not 
perform as well as WA-DPDS.

Similar results were achieved for ensembles consisting of workflows of equal
size set to 1000. A summary of these results is shown in
Table~\ref{tab:num-dags-equal}. Again, we can see that the static SPSS algorithm performed better for coarse-grained
workflows (LIGO, Epigenomics, SIPHT). The differences between algorithm scores
are smaller since the workflows have equal size, therefore rejecting a particular
workflow does not make it any easier for the algorithms to complete any other 
workflow.

Figure~\ref{fig:budgets} shows the same data for the SIPHT ensemble from another
perspective. It helps answer the question: Given a deadline, how many workflows
can be finished when we increase the budget? In Figure~\ref{fig:budgets}(a) the
growth is super-linear, which results from the Pareto-like distribution of 
workflow sizes in the ensemble (see Figure~\ref{fig:ensemble-pareto}). 
For ensembles of constant size workflows (see Figure~\ref{fig:budgets}(b)) 
the growth is linear and the costs are higher, since all the workflows in the
test consisted of 1000 tasks.

\begin{figure}[htb]
    \centering
    \subfloat[size]{
        \includegraphics[width=0.75\columnwidth]{size-selected-LIGO}
        }\\
    \subfloat[cost]{
        \includegraphics[width=0.75\columnwidth]{cost-selectedLIGO}
        }
    \caption{Example performance of algorithms expressed in
    sum of dag sizes and effective cost, data for LIGO application, Pareto
    Sorted ensemble of 100 workflows.}
    \label{fig:size-cost}
\end{figure}

>>>>

\begin{figure}[tb] 
  \centering
  \includegraphics[width=0.5\textwidth]{figures/spss-gantt}
  \caption[Example schedule generated by the SPSS algorithm]{Example schedule generated by the SPSS algorithm. Each row is a different VM. Tasks are represented as boxes and are colored by workflow.}
  \label{fig:spss-example}
\end{figure}

@@

\subsection{Performance Metrics}
\label{sec:perf_metric}

@@

@@

@@

\TODO{Edit this first paragraph, which was moved from the old Scenarios section}

@@

\begin{figure*}[htb] 
\TODO{Cut this figure down}\\
\centering
\includegraphics[width=0.19\textwidth]{pareto-cost-MONTAGE-n-1000-8-dagh1-20m0}
\includegraphics[width=0.19\textwidth]{pareto-cost-CYBERSHAKE-n-1000-8-dagh1-20m0}
\includegraphics[width=0.19\textwidth]{pareto-cost-LIGO-n-1000-8-dagh1-40m0}
\includegraphics[width=0.19\textwidth]{pareto-cost-GENOME-n-1000-8-dagh100-1500m0}
\includegraphics[width=0.19\textwidth]{pareto-cost-SIPHT-n-1000-8-dagh5-50m0}
\caption{Effective computing cost in \$ per hour calculated by dividing the sum of
task runtimes by the total budget. Higher costs result from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows. The ranges on $y$ axis are chosen to 
emphasize the differences  between  the performance of the algorithms.}
\label{fig:cost}
\end{figure*}

\TODO{Rewrite this next paragraph}

@@

\begin{figure}[htb]
    \TODO{Does this figure need removed?}
    \centering
    \subfloat[size]{
        \includegraphics[width=0.75\columnwidth]{size-selected-LIGO}
        }\\
    \subfloat[cost]{
        \includegraphics[width=0.75\columnwidth]{cost-selectedLIGO}
        }
    \caption{Example performance of algorithms expressed in
    sum of dag sizes and effective cost, data for LIGO application, Pareto
    Sorted ensemble of 100 workflows.}
    \label{fig:size-cost}
\end{figure}
